---
title: "Kay_Shannon_APSTA_2012_Final_Project"
author: "Shannon Kay"
date: "12/9/2019"
output: html_document
---

```{r}
##load required packages
library(tidyverse)
library(AER)
library(MatchIt)

```

How do post-treatment variables affect estimates of treatment effects? 

Step 1: Construct our data

Imagine a group of 1000 NYU students are randomly encouraged to attend an optional, weekly lab to supplement their statistics class. Approximately half of the students enrolled in Stats 101 receive the information regarding when and where the lab meets. Students who do not receive the information may attend, but would have to discover the

```{r}
set.seed(2012)

compliance_type <- rep(c("c","n", "a"), c(500, 300, 200))

d_0 <- rep(c(0,0,1), c(500, 300, 200))
d_1 <- rep(c(1,0,1), c(500, 300, 200))

dat<- data.frame(d_0, d_1, compliance_type)

```

Determine what baseline levels are appropriate for each group, then add treatment effect, & create both possible observed outcomes 

(c) Simulate the potential outcomes in a way that meets the following criteria:
  (i) The exclusion restriction is satisfied.
  (ii) The average effect of Z on Y for the compliers is 9.
  (iii) The average Y(Z=0) for never takers is 0; The average Y(0) for compliers is 3; The average Y(Z=0) for always takers is 6.
  (iii) The residual standard deviation is 1 for everyone in the sample (generated independently for each potential outcome).
  
```{r}

comp_Y0 <- rnorm(500, 80, 3.5)
nev.taker_Y0 <- rnorm(300,75,7)
always_Y0 <- rnorm(200,87,3.5)

dat$Y0 <- c(comp_Y0, nev.taker_Y0, always_Y0)
hist(dat$Y0)
range(dat$Y0)

comp_Y1 <- rnorm(500,89,3.5)
nev.taker_Y1 <- rnorm(300,75,7)
always_Y1 <- rnorm(200,87,3.5)

dat$Y1 <- c(comp_Y1, nev.taker_Y1, always_Y1)

hist(dat$Y1)
range(dat$Y1)

dat$age <- rnorm(1000,20,.5)

dat$gender <- rbinom(1000, size = 1, prob = .5)

##checking sate for compliers
sate_compliers <- mean(comp_Y1) - mean(comp_Y0)
```


## Question 2. Playing the role of the researcher to randomly assign treatments to observations.
Now switch to the role of the researcher. Pretend that you are running the experiment that we are examining for this assignment.  Generate a binary indicator for the ignorable treatment *assignment* (as distinct from treatment receipt.... so this is $Z$, not $D$).  Probability of being assigned to the treatment condition should be .5.

```{r}

dat.full <- dat

dat.full$Z <- rbinom(1000, size = 1, prob = .5)

```


##Create dosage variables

```{r}

## % attendance is always the same for compliars not encouraged, never-takers, & always takers
comp_attend_Z0 <- rep(0,500)
nev.taker_attend <- rep(0,300)
always_attend <- rep(1,200)

##varying attendance for compliers
## % of attendance for compliers skewed lower- students attend once or twice and dwindle off
comp_attend1_Z1 <- rbeta(500,shape1 = 2, shape2 = 15)

## % of attendance for compliers  
comp_attend2_Z1 <- rbeta(500,shape1 = 2, shape2 = 2)

## % attendance for compliers 
comp_attend3_Z1 <- rbeta(500,shape1 = 2, shape2 = .5)

dat.full$attend_Z0 <- c(comp_attend_Z0, nev.taker_attend, always_attend)

dat.full$attend1_Z1 <- c(comp_attend1_Z1, nev.taker_attend, always_attend)
dat.full$attend2_Z1 <- c(comp_attend2_Z1, nev.taker_attend, always_attend)
dat.full$attend3_Z1 <- c(comp_attend3_Z1, nev.taker_attend, always_attend)


```

##Create observed data
```{r}

Y_obs <- ifelse(dat.full$Z== 1, dat.full$Y1, dat.full$Y0)
D_obs <- ifelse(dat.full$Z== 1, dat.full$d_1, dat.full$d_0)

attend1_obs <- ifelse(dat.full$Z== 1, dat.full$attend1_Z1, dat.full$attend_Z0)
attend2_obs <- ifelse(dat.full$Z== 1, dat.full$attend2_Z1, dat.full$attend_Z0)
attend3_obs <- ifelse(dat.full$Z== 1, dat.full$attend3_Z1, dat.full$attend_Z0)

##pull alll observed variables into the observed data frame
##keeping some extra variables to look at correlation
dat.obs <- data.frame(dat.full$Z, compliance_type, D_obs, Y_obs, dat.full$gender, dat.full$age, attend1_obs, attend2_obs, attend3_obs)

##rename columns
colnames(dat.obs) <- c("Z", "Compliance", "D", "Y", "Gender", "Age", "Attend_1", "Attend_2", "Attend_3")

##looking at correlation to compliance & to treatment
kruskal.test(attend1_obs ~ Compliance, data = dat.obs)
kruskal.test(attend2_obs ~ Compliance, data = dat.obs)
kruskal.test(attend3_obs ~ Compliance, data = dat.obs)

cor(dat.obs$Y, dat.obs$Attend_1)
cor(dat.obs$Y, dat.obs$Attend_2)
cor(dat.obs$Y, dat.obs$Attend_3)

plot(hist(dat.obs$Attend_1[dat.obs$Compliance=="c"]))
plot(hist(dat.obs$Attend_2[dat.obs$Compliance=="c"]))
plot(hist(dat.obs$Attend_3[dat.obs$Compliance=="c"]))


```


```{r}

naive_regression <- lm(Y ~ D, data = dat.obs)
summary(naive_regression)

naive_regression_w.predictors <- lm(Y ~ D + Age + as.factor(Gender), data = dat.obs)
summary(naive_regression_w.predictors)

naive_regression_w.attend_2 <- lm(Y ~ Z + Age + as.factor(Gender) + Attend_1, data = dat.obs)
summary(naive_regression_w.attend_2)

naive_regression_w.attend_3 <- lm(Y ~ D + Age + as.factor(Gender) + Attend_3, data = dat.obs)
summary(naive_regression_w.attend_3)


##In dat.obs, 21.88% are always-takers, 48.43% are compliers, and 29.68% are never-takers.

compliance_regression <- lm(D ~ Z, data = dat.obs)

summary(compliance_regression)

1-(compliance_regression$coefficients[1]+compliance_regression$coefficients[2])

```

Intent to Treat Effect
ITT = 4.38
CACE = 9.05
```{r}

ITT <- mean(dat.obs$Y[dat.obs$Z == 1]) - mean(dat.obs$Y[dat.obs$Z ==0])
ITT

## Estimate the CACE by dividing the ITT estimate by the percent of compliers in the sample.

CACE <- ITT/compliance_regression$coefficients[2]
CACE

```

Provide an estimate of CACE and its standard error using the ivreg command in the AER package.

```{r}

ivreg.fit <- ivreg(Y ~ D | Z, data = dat.obs)
ivreg_summary <- summary(ivreg.fit)
ivreg_summary

CACE_ivreg <- ivreg_summary$coefficients[2]
std.error_ivreg <- ivreg_summary$coefficients[,2]

CACE_ivreg
std.error_ivreg
##number below D for std. error
```


```{r}
set.seed(2012)

data_function1 <- function(x){
  c <- rep(c("c", "n", "a"), x*c(.5,.3,.2))
  d_0 <- rep(c(0,0,1), x*c(.5, .3, .2))
  d_1 <- rep(c(1,0,1), x*c(.5, .3, .2))
  
dat<- data.frame(d_0, d_1, compliance_type)

  comp_Y0 <- rnorm(x*.5, 80, 3.5)
  nev.taker_Y0 <- rnorm(x*.3,75,7)
  always_Y0 <- rnorm(x*.2,87,3.5)

  Y0 <- c(comp_Y0, nev.taker_Y0, always_Y0)

  comp_Y1 <- rnorm(x*.5,89,3.5)
  nev.taker_Y1 <- rnorm(x*.3,75,7)
  always_Y1 <- rnorm(x*.2,87,3.5)

  Y1 <- c(comp_Y1, nev.taker_Y1, always_Y1)
  
  age <- rnorm(1000,20,.5)
  gender <- rbinom(1000, size = 1, prob = .5)
  
  comp_attend_Z0 <- rep(0,500)
  nev.taker_attend <- rep(0,300)
  always_attend <- rep(1,200)

  ##varying attendance for compliers
  ## % of attendance for compliers skewed lower- students attend once or twice and dwindle off
  comp_attend1_Z1 <- rbeta(500,shape1 = 2, shape2 = 15)
  ## % of attendance for compliers  
  comp_attend2_Z1 <- rbeta(500,shape1 = 2, shape2 = 2)
  ## % attendance for compliers 
  comp_attend3_Z1 <- rbeta(500,shape1 = 2, shape2 = .5)

  attend_Z0 <- c(comp_attend_Z0, nev.taker_attend, always_attend)
  attend1_Z1 <- c(comp_attend1_Z1, nev.taker_attend, always_attend)
  attend2_Z1 <- c(comp_attend2_Z1, nev.taker_attend, always_attend)
  attend3_Z1 <- c(comp_attend3_Z1, nev.taker_attend, always_attend)
  
  Z <- rbinom(x, size = 1, prob = .5)

  god.dat <- data.frame(d_0, d_1, c, Y0, Y1, age, gender, attend1_Z1, attend2_Z1, attend3_Z1, Z)
  
  Y_obs <- ifelse(god.dat$Z== 1, god.dat$Y1, god.dat$Y0)
  D_obs <- ifelse(god.dat$Z == 1, god.dat$d_1, god.dat$d_0)
  
  attend1_obs <- ifelse(dat.full$Z== 1, dat.full$attend1_Z1, dat.full$attend_Z0)
  attend2_obs <- ifelse(dat.full$Z== 1, dat.full$attend2_Z1, dat.full$attend_Z0)
  attend3_obs <- ifelse(dat.full$Z== 1, dat.full$attend3_Z1, dat.full$attend_Z0)

  sim.dat.obs <- data.frame(D_obs, god.dat$Z, god.dat$age, god.dat$gender, attend1_obs, attend2_obs, attend3_obs, Y_obs)
  colnames(sim.dat.obs) <- c("D", "Z", "Age", "Gender", "Attend1", "Attend2", "Attend3", "Y")
  
  return(sim.dat.obs)
}


##set # draws, initialize vector to store CACE estimates
draws <- 10000
CACE_ivreg_baseline <- numeric()
CACE_ivreg_2 <- numeric()
CACE_ivreg_3 <- numeric()
CACE_ivreg_4 <- numeric()

for(i in 1:draws){
  sim.dat <- data_function(1000)
  
  ##baseline function
  ivreg.baseline <- ivreg(Y ~ D + age + gender | Z, data = sim.dat)
  CACE_ivreg_baseline[length(CACE_ivreg_baseline)+1] <- ivreg.baseline$coefficients[2]
  
  ##fit 2
  ivreg.fit2 <- ivreg(Y ~ D + age + gender + Attend1 | Z, data = sim.dat)
  CACE_ivreg_2[length(CACE_ivreg_2)+1] <- ivreg.fit2$coefficients[2]
  
  ##fit 3
  ivreg.fit3 <- ivreg(Y ~ D + age + gender + Attend2 | Z, data = sim.dat)
  CACE_ivreg_3[length(CACE_ivreg_3)+1] <- ivreg.fit3$coefficients[2]
  
  ##fit 4
  ivreg.fit4 <- ivreg(Y ~ D + age + gender + Attend3 | Z, data = sim.dat)
  CACE_ivreg_4[length(CACE_ivreg_4)+1] <- ivreg.fit4$coefficients[2]
  
}

sim_CACE_std.dev <- sd(CACE_ivreg_sim)
sim_CACE_std.dev

mean_CACE_sim <- mean(CACE_ivreg_sim)
mean_CACE_sim

```
