---
title: "Kay_Shannon_APSTA_2012_Final_Project"
author: "Shannon Kay"
date: "12/9/2019"
output: html_document
---

```{r}
##load required packages
library(tidyverse)


```


How do post-treatment variables affect estimates of treatment effects? 

Step 1: Construct our data

Imagine a group of 1000 incoming NYU freshmen are randomly encouraged to attend an optional, weekly lab to supplement their statistics class. Approximately half of the students enrolled in Stats 101 receive the information regarding when and where the lab meets. Students who do not receive the information may attend, but would have to discover the

```{r}

compliance_type <- rep(c("c","n", "a"), c(450, 325, 225))

d_0 <- rep(c(0,0,1), c(450, 325, 225))
d_1 <- rep(c(1,0,1), c(450, 325, 225))

dat<- data.frame(d_0, d_1, compliance_type)


```

Determine what baseline levels are appropriate for each group, then add treatment effect, & create both possible observed outcomes 

(c) Simulate the potential outcomes in a way that meets the following criteria:
  (i) The exclusion restriction is satisfied.
  (ii) The average effect of Z on Y for the compliers is 4.
  (iii) The average Y(Z=0) for never takers is 0; The average Y(0) for compliers is 3; The average Y(Z=0) for always takers is 6.
  (iii) The residual standard deviation is 1 for everyone in the sample (generated independently for each potential outcome).
  
```{r}

comp_Y0 <- rnorm(450, 80, 3.5)
nev.taker_Y0 <- rnorm(325,75,7)
always_Y0 <- rnorm(225,87,3.5)

hist(always_Y0)
range(always_Y0)

dat$Y0 <- c(comp_Y0, nev.taker_Y0, always_Y0)
hist(dat$Y0)

comp_Y1 <- rnorm(450,90,3.5)
nev.taker_Y1 <- rnorm(325,75,7)
always_Y1 <- rnorm(225,87,3.5)

dat$Y1 <- c(comp_Y1, nev.taker_Y1, always_Y1)

hist(dat$Y1)
range(dat$Y1)


##checking sate for compliers
sate_compliers <- mean(comp_Y1) - mean(comp_Y0)
```


## Question 2. Playing the role of the researcher to randomly assign treatments to observations.
Now switch to the role of the researcher. Pretend that you are running the experiment that we are examining for this assignment.  Generate a binary indicator for the ignorable treatment *assignment* (as distinct from treatment receipt.... so this is $Z$, not $D$).  Probability of being assigned to the treatment condition should be .5.

```{r}

dat.full <- dat

dat.full$Z <- rbinom(1000, size = 1, prob = .5)

```


##Create dosage variables

```{r}

comp_attend1_X0 <- rep(0,450)
nev.taker_attend1_X0 <- rep(0,325)
always_attend1_X0 <- rep(1,225)

dat.full$attend1_X0 <- c(comp_attend1_X0, nev.taker_attend1_X0, always_attend1_X0)

comp_attend1_X1 <- rep(1,450)
nev.taker_attend1_X1 <- rep(0,325)
always_attend1_X1 <- rep(1,225)

dat.full$attend1_X1 <- c(comp_attend1_X1, nev.taker_attend1_X1, always_attend1_X1)

```



```{r}

#cor(dat.obs$Z, dat.obs$Attend_2)

##violates compliers.....
#dat.obs$Attend_2 <- rbinom(1000, size = 1, prob = .5)




```

Varying levels of dosage

```{r}

attendance <- c(1:10)

comp_attend2_X0 <- rep(0,450)
nev.taker_attend2_X0 <- rep(0,325)
always_attend2_X0 <- rep(1,225)

dat.full$attend2_X0 <- c(comp_attend2_X0, nev.taker_attend2_X0, always_attend2_X0)

comp_attend2_X1 <- sample(attendance,450, replace = TRUE)
nev.taker_attend2_X1 <- rep(0,325)
always_attend2_X1 <- rep(1,225)

dat.full$attend2_X1 <- c(comp_attend2_X1, nev.taker_attend2_X1, always_attend2_X1)



```


##Create observed data

```{r}

Y_obs <- ifelse(dat.full$Z== 1, dat.full$Y1, dat.full$Y0)

D_obs <- ifelse(dat.full$Z== 1, dat.full$d_1, dat.full$d_0)

attend1_obs <- ifelse(dat.full$Z== 1, dat.full$attend1_X1, dat.full$attend1_X0)

attend2_obs <- ifelse(dat.full$Z == 1, dat.full$attend2_X1, dat.full$attend2_X0)

dat.obs <- data.frame(dat.full$Z, D_obs, Y_obs, attend1_obs, attend2_obs, compliance_type)

colnames(dat.obs) <- c("Z", "D", "Y", "Attend_1", "Attend_2", "compliance")

kruskal.test(attend1_obs ~ compliance, data = dat.obs)
kruskal.test(attend2_obs ~ compliance, data = dat.obs)

hist(dat.obs$Y)

```


